---
title: "Analyzing Social Vulnerability Indicators for Potential Coastal Hazard in Hawai'i"
format: html
editor: visual
editor_options: 
  chunk_output_type: console
---

Load in Libraries

```{r}
library("tidyverse")
library("here")
library("sf")
library("mapview")
library("mapgl")
library("ggiraph")
library("leaflet")
library("shiny")
```

Read in Data

```{r}
block_indicators <- st_read("data/2020_Social_Vulnerability_Indicators_for_Coastal_Hazards__by_Block_Group.geojson")

coastal_hazards <- st_read("data/SLR_Exposure_Area_-_3.2_Ft._Scenario.geojson")
```

Education: Percent Population over the age of 25 with no high-school diploma in the bg
Vehicles: Percent Households with no access to a vehicle in the bg
Disability: Percent Population with a disability in the bg
Over_65: Percent Population over the age of 65 in the bg
Under_5: Percent Population under the age of 5 in the bg
Single_Par: Percent Single Parent Households w/ children under the age of 18 in the bg
Household: Percent Households with no internet access or limited internet access and/or
no computer in the bg
Limited_En: Percent Households that speak "limited English” in the bg
Built_befo: Percent Households built before 1970 in the bg
Housing_Te: Percent Households that are renter-occupied in the bg
Industry_A: Percent Population over the age of 16 employed in nature based or service
industries in the bg
Househol_1: Percent Population of female householders in the bg
Hispanic_L: Percent Population that identifies as Hispanic or Latino in the bg
Race: Percent Population that is Native Hawaiian or Pacific Islander in the bg
Poverty: Percent Population for whom poverty status is determined in the bg
Income_Def: the Aggregate income deficit (dollars) in the past 12 months for families within
the given block group
Count: Total Number of Social Vulnerability Indicators that are true for this block
group

## Changing data types in DF as int or dbls(floating numbers)

```{r}
block_indicators <- block_indicators %>%
  mutate(across(c(education, vehicles, disability, overr_65, under_5,
                   single_par, household, limited_en, built_befo,
                   housing_te, industry_a, househol_1, hispanic_l,
                   race, poverty), as.numeric))

```

Chlorpleth Map

```{r}

pal <- colorNumeric(palette = "YlOrRd", domain = block_indicators$count_, na.color = "transparent")

leaflet(block_indicators) %>%
  addProviderTiles("CartoDB.Positron") %>%
  addPolygons(
    fillColor = ~pal(count_),
    fillOpacity = 0.7,
    color = "white",
    weight = 0.5,
    label = ~paste0("Vulnerability Count: ", count_),
    highlightOptions = highlightOptions(weight = 2, color = "black", bringToFront = TRUE)
  ) %>%
  addLegend(pal = pal, values = ~count_, title = "Vulnerability Count", position = "bottomright")
```

Lets see what our 2nd dataset looks like on a map

```{r}
mapview(coastal_hazards)

```

Changes to our data frame

-   re-adjust data types in dataframe to int or dbl

-   remove irrelevant vulnerabilities to coastal disaster

-   re-calculate our count totals

```{r}
block_indicators <- block_indicators %>%
  mutate(across(c(education, vehicles, disability, overr_65, under_5,
                   single_par, household, limited_en, built_befo,
                   housing_te, industry_a, househol_1, hispanic_l,
                   race, poverty), as.numeric))

block_less <- block_indicators %>%
  select(-objectid, -bg, -industry_a, -hispanic_l, -race, -st_perimetershape, -st_areashape)

block_less <- block_less %>%
  mutate(count_ = rowSums(!is.na(across(c(education, vehicles, disability, overr_65,
                                            under_5, single_par, household, limited_en,
                                            built_befo, housing_te, househol_1, poverty))))
                 + (income_def != 0))
```

Updated map with 2 layers: vulnerability variables by area block group and projected coastal disaster locations

```{r}
pal <- colorNumeric(palette = "YlOrRd", domain = block_less$count_, na.color = "transparent")

leaflet(block_less) %>%
  addProviderTiles("CartoDB.Positron") %>%
  addPolygons(
    fillColor = ~pal(count_),
    fillOpacity = 0.7,
    color = "white",
    weight = 0.5,
    label = ~paste0("Vulnerability Count: ", count_, " | Population: ", pop20),
    highlightOptions = highlightOptions(weight = 2, color = "black", bringToFront = FALSE),
    group = "Vulnerability",
  ) %>%
  addPolygons(
    data = coastal_hazards,
    fillColor = "blue",
    fillOpacity = 0.4,
    color = "blue",
    weight = 1,
    group = "Coastal Hazard Projection Zones (3.2 ft)"
  ) %>%
  addLegend(pal = pal, values = ~count_, title = "Vulnerability Count", position = "bottomright") %>%
  addLayersControl(
    overlayGroups = c("Vulnerability", "Coastal Hazard Projection Zones (3.2 ft"),
    options = layersControlOptions(collapsed = FALSE)
  )
```

Shiny APP (I couldn't upload deploy the app, I kept getting the same error) - made a HTML instead and put it on git

```{r}
ui <- fluidPage(
  titlePanel("Analyzing Social Vulnerability Indicators for Potential Coastal Hazard in Hawai'i"),
  mainPanel(
    leafletOutput("map", height = "700px")
  )
)

server <- function(input, output, session) {
  output$map <- renderLeaflet({
    pal <- colorNumeric(palette = "YlOrRd", domain = block_less$count_, na.color = "transparent")

    leaflet(block_less) %>%
      addProviderTiles("CartoDB.Positron") %>%
      addPolygons(
        fillColor = ~pal(count_),
        fillOpacity = 0.7,
        color = "white",
        weight = 0.5,
        label = ~paste0("Vulnerability Count: ", count_, " | Population: ", pop20),
        highlightOptions = highlightOptions(weight = 2, color = "black", bringToFront = FALSE),
        group = "Vulnerability"
      ) %>%
      addPolygons(
        data = coastal_hazards,
        fillColor = "blue",
        fillOpacity = 0.4,
        color = "blue",
        weight = 1,
        group = "Coastal Hazard Projection Zones (3.2 ft)"
      ) %>%
      addLegend(pal = pal, values = ~count_, title = "Vulnerability Count", position = "bottomright") %>%
      addLayersControl(
        overlayGroups = c("Vulnerability", "Coastal Hazard Projection Zones (3.2 ft)"),
        options = layersControlOptions(collapsed = FALSE)
      )
  })
}

shinyApp(ui, server)
```
